Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-10 paper/ijm.tex
Reacted to several comments by adding new comments. Copy and pasted some lines to start the CEE comparison part. Removed the $g$ index from  $\upsilon_{t}$.

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.

@@

We describe the outcome of individual $i$ belonging to group $g$ (either Estonian
or Russian) at year $t$ as
\begin{equation}
	\label{eq:specification}
	\begin{split}
		y_{it} ^{g} = &\: \alpha^{g}{}' \vec{X}_{it}^{g} + \vec{\beta}^{g}{}' \cdot \vec{L}_{it}^{g} +
		\eta_{i}^{g} +
		\upsilon_{t}^{g} + \epsilon_{it}^{g},
	\end{split}
\end{equation}
where $y$ is the outcome (either unemployment indicator
or logarithm of net monthly wage), $\vec{X}$ are the other individual
characteristics; $\vec{L}$ is the vector of
language skill descriptors, $\eta_{i}^{g}$ is the individual-specific effect, $\upsilon_{t}^{g}$ are the year dummies and $\epsilon_{it}^{g}$
is the idiosyncratic error term. 
\fxerror{ot: what's the eta term?  do you do random effects?}
\fxerror{ot: why does $\upsilon$ have index $g$?}
As we use panel data, we cluster the
standard errors on individuals.  When analysing the long-term time
trend, we use the same specification (without year fixed effects) to
estimate separate models for each individual year.

We estimate three models that differ in terms of the number of
control variables. The first specification only includes the language skills
and year dummies.  It describes the ``raw'' effect of language skills
on outcomes after controlling for substantial time trends in our data.
The second specification additionally includes individual
characteristics (age, education) and family descriptors (marital
status, indicator for interethnic marriage, having kids up to 17 years
old) and county of residence.  
Finally, the third specification also adds industry, occupation, and other
workplace descriptors.
Note that in case of unemployment probability,
we only estimate the first two models.

@@

The most surprising result in our analysis so far--the fact that
Russian men earn virtually no Estonian language premium while women do--warrants a
deeper look at the data.  We compute the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills, residence in Tallinn metro area,
and education.  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from that in the North-East.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In addition, to correct very strong wage and language skill trends through these
years, we also present the average
residuals from an OLS regression that explains log wage by year fixed
effects (Table~\ref{tab:loc-lang-edu}). 
\fxwarning{ot: Sven, can you redo this table?  Use exactly the same
  sample as you used.  Would also be good to put both et1 and et0 on
  the same row}
The table shows that the picture is in fact quite complex.  We can see
that there is little to no gain from Estonian skills for the
largest group, high-school educated workers.  Even more, outside of
the metropolitan area, the lowest and highest education groups
have in fact negative language premium.  However, these two groups
enjoy quite a noticeable premium in the Tallinn region.  The
opposite-signed effects in for these two education groups suggests
that a
selective internal migration may play a role.

@@

In summary, this dataset supports the previous hypothesis: Russian women are
working in less segregated environment where Estonian is the main
communication language, and they need Estonian at work for other reasons as
well.  Gender differences in English usage are not statistically
significant.
\fxwarning{ot: let's remove the cv.ee stuff.  Or substantially improve
that.}

>>>>

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
\fxwarning{SKB: Let's collect first enough material and then decide, but I can start adding here some notes from \cite{Kahanec2009,Kahanec2010}.}
%Start here CEE stuff for now
\cite{Kahanec2009} provide an overview about \enquote{the roles of foreign origin and citizenship in the labor markets of Eastern and Eestern[sic] Member States of the EU} \dots and use the EU Survey of Income and Living Conditions (EU-SILC) dataset.
They\enquote{ mapped and compared the roles of foreign origin and citizenship for labor market performance in Eastern and Western EU Member States.}
Using Oaxaca and Fairlie decompositions, they show that there is native/non-native divide in the national labour markets \dots.
They find that \enquote{both foreign origin and citizenship remain important barriers to non-native labor market integration}
\enquote{In Estonia and Latvia, in particular, employment
	policies seem to be most relevant for non-immigrant non-citizens. Once employed,
	non-nativity may play a less negative role in the labor market for some groups,
	although labor market segmentation remains a significant issue.}
\fxnote{\cite{Kahanec2009} essentially confirms your results in \cite{Leping2008,Toomet2011}}
\cite{Kahanec2010} show that \enquote{the labor market situation of ethnic minorities in the European
	Union. Facing a serious measurement challenge and lacking adequate data, we apply
	several measures of ethnicity and examine various data sources as well as secondary
	evidence. We find significant gaps between ethnic minority and majority populations in
	terms of labor market outcomes.}
%Continue with existing stuff
In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.

@@

We describe the outcome of individual $i$ belonging to group $g$ (either Estonian
or Russian) at year $t$ as
\begin{equation}
	\label{eq:specification}
	\begin{split}
		y_{it} ^{g} = &\: \alpha^{g}{}' \vec{X}_{it}^{g} + \vec{\beta}^{g}{}' \cdot \vec{L}_{it}^{g} +
		\eta_{i}^{g} +
		\upsilon_{t}^{g} + \epsilon_{it}^{g},
	\end{split}
\end{equation}
where $y$ is the outcome (either unemployment indicator
or logarithm of net monthly wage), $\vec{X}$ are the other individual
characteristics; $\vec{L}$ is the vector of
language skill descriptors, $\eta_{i}^{g}$ is the individual-specific effect, $\upsilon_{t}$ are the year dummies and $\epsilon_{it}^{g}$
is the idiosyncratic error term. 
\fxerror{ot: what's the eta term?  do you do random effects?}
\fxerror{SKB: Eta should be the individual specific random effect. We cluster the standard errors at the individual level, hence estimate a random-effect model, although not a panel random-effect model.}
\fxerror{ot: why does $\upsilon$ have index $g$?}
\fxerror{SKB: Should be an error, probably index $g$ was just added to everything.}
As we use panel data, we cluster the
standard errors on individuals.  When analysing the long-term time
trend, we use the same specification (without year fixed effects) to
estimate separate models for each individual year.

We estimate three models that differ in terms of the number of
control variables. The first specification only includes the language skills
and year dummies.  It describes the ``raw'' effect of language skills
on outcomes after controlling for substantial time trends in our data.
The second specification additionally includes individual
characteristics (age, education) and family descriptors (marital
status, indicator for inter-ethnic marriage, having kids up to 17 years
old) and county of residence.  
Finally, the third specification also adds industry, occupation, and other
workplace descriptors.
Note that in case of unemployment probability,
we only estimate the first two models.
\fxnote{Actually, I tried to estimate Model 3 as well for unemployment probability. The language coefficients become insignificant after controlling for industry and occupation. Hence, Estonian language fluency lowers the unemployment probability only if workplace characteristics are ignored. Since nobody questioned this result during review, we should maybe not mention it now. }

@@

The most surprising result in our analysis so far--the fact that
Russian men earn virtually no Estonian language premium while women do--warrants a
deeper look at the data.  We compute the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills, residence in Tallinn metro area,
and education.  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from that in the North-East.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In addition, to correct very strong wage and language skill trends through these
years, we also present the average
residuals from an OLS regression that explains log wage by year fixed
effects (Table~\ref{tab:loc-lang-edu}). 
\fxwarning{ot: Sven, can you redo this table?  Use exactly the same
  sample as you used.  Would also be good to put both et1 and et0 on
  the same row}
\fxwarning{SKB: Will try. Although I still do not understand, what the residual is telling me and why the wages are so high. Are the wages in EEK? What does ET stand for? The table needs more explanation or at least a proper legend.}
The table shows that the picture is in fact quite complex.  We can see
that there is little to no gain from Estonian skills for the
largest group, high-school educated workers.  Even more, outside of
the metropolitan area, the lowest and highest education groups
have in fact negative language premium.  However, these two groups
enjoy quite a noticeable premium in the Tallinn region.  The
opposite-signed effects in for these two education groups suggests
that a
selective internal migration may play a role.

@@

In summary, this dataset supports the previous hypothesis: Russian women are
working in less segregated environment where Estonian is the main
communication language, and they need Estonian at work for other reasons as
well.  Gender differences in English usage are not statistically
significant.
\fxwarning{ot: let's remove the cv.ee stuff.  Or substantially improve
that.}
\fxwarning{SKB: You removed the cv.ee already? I only see stuff using data from Integration Monitor Survey 2008. I did not do these analyses nor do I have the data to repeat the analyses for the latest edition of the Integration Monitor Survey.
The cv.ee might have been interesting or useful, if it was done properly documented. But that might warrant another section or even another paper. See \cite{fabo+2017E} for more on that.
}
